---
title: "new_labels"
output: pdf_document
date: "2025-05-03"
---




## Load data
```{r}

metadata <- read.csv("DataSets/2_analysis/metadata.csv", row.names = 1)

survival <- read.csv("DataSets/2_analysis/survival_data.csv", row.names = 1)

mutations <- read.csv("DataSets/2_analysis/mutations.csv", row.names = 1)
dna <- read.csv("DataSets/2_analysis/dna.csv", row.names = 1)
mrna <- read.csv("DataSets/2_analysis/mrna.csv", row.names = 1)
mirna <- read.csv("DataSets/2_analysis/mirna.csv", row.names = 1)


info.methy <- read.csv("DataSets/processed_assays/info_methylation.csv", row.names = 1)
info.mirna <- read.csv("DataSets/processed_assays/info_rna_mirna.csv", row.names = 1)
info.mrna <- read.csv("DataSets/processed_assays/info_rna_coding.csv", row.names = 1)
```


```{r}
omics.list <- list(Mutations = mutations, Methylation = dna, mRNA = mrna, miRNA = mirna)
```


```{r}
factors_mofa <- as.matrix(read.csv("MOFA/output/factors_mofa.csv", row.names = 1))

loadings_mofa <- list(
  Mutations = as.matrix(read.csv("MOFA/output/loadings_mofa_mut.csv", row.names = 1)),
  Methylation = as.matrix(read.csv("MOFA/output/loadings_mofa_dna.csv", row.names = 1)),
  mRNA = as.matrix(read.csv("MOFA/output/loadings_mofa_mrna.csv", row.names = 1)),
  miRNA = as.matrix(read.csv("MOFA/output/loadings_mofa_mirna.csv", row.names = 1))
)
```


```{r}
survival_omics <- as.matrix(read.csv("DataSets/2_analysis/survival_data.csv", row.names = 1))
```



## Analysis

```{r}
library(survival)
library(survminer)
```

# Find the best number of clusters (3,4,5,6), based on p-value of log-rank test

```{r}
library("dplyr")

res_l <- list()
clust_l <- list()
for (k in 3: 8) {  
  
set.seed(4)  # for reproducibility
 
  res <- c()

  for (r in 1: 30) {
    strat_sample <- survival_omics %>%
      tibble::rownames_to_column("SampleID") %>%
      group_by(Status) %>%
      sample_frac(size = 0.80) %>%
      tibble::column_to_rownames("SampleID")


    km_res <- kmeans(factors_mofa[rownames(strat_sample), c(1,2,3)], centers = k, nstart = 10)
    
    
    surv_obj <- Surv(time = strat_sample[, "Time"] , 
                     event = strat_sample[, "Status"])
  
    fit <- survfit(surv_obj ~ factor(km_res$cluster))
    
    logrank <- survdiff(surv_obj ~ factor(km_res$cluster))
    
    p_val <- logrank$pvalue
    
    res <- c(res, p_val)
    
  }
  res_l[[as.character(k)]] <- res
}
```



```{r}
res_l_log10 <- sapply(res_l, function(x) -log10(x))


df <- data.frame(res_l_log10)
colnames(df) <- c(3,4,5,6,7,8)

df_long <- df %>%
  pivot_longer(
    cols = everything()
  )



p <- ggplot(df_long, aes(x = name, y = value, fill = name)) +
  geom_boxplot(fill = "#FF7F32") +
  labs(x = "Number of Clusters", y = "-log10(p-value)") +
  theme(
      axis.text.y = element_text(size=rel(text_size), hjust=1, color='black'),
      axis.text.x = element_text(size=rel(text_size), vjust=0.5, color='black'),
      axis.title.y=element_text(size=rel(text_size), hjust=1, color='black'),
      axis.title.x=element_text(size=rel(text_size), vjust=1, color='black'),
      legend.position="none",
      legend.text =  element_blank(), 
      legend.title =  element_blank(), ,
      panel.grid.minor = element_blank(), 
      panel.border = element_rect(color = "black", fill = NA, size = 0.8),
      panel.background = element_blank(),
    ) 


ggsave(filename = paste0("search_labels/", "dist_logp_k", ".pdf"), plot = p, width = 10, height = 7, dpi = 300)
```



## K = 5

# Plot the projection based on that assignment + Survival Curves of each cluster

```{r}

k <- 5

set.seed(44)
km_res <- kmeans(factors_mofa[,c(1,2,3)], centers = k, nstart = 10)

```

```{r}

plot_proj_data <- data.frame(x = factors_mofa[,1], y  = factors_mofa[,3], color_by = factor(km_res$cluster,levels = 1:5, 
                                  labels = paste("Cluster", 1:5)))


my_colors <- brewer.pal(n = 5, name = "Dark2")
p <- ggplot(plot_proj_data , aes(x=.data$x, y=.data$y, color=.data$color_by)) + 
    geom_point(size=3.0) +
    labs(x="Factor 1", y="Factor 3", color = "") +
    scale_color_manual(values = my_colors) +
     theme(
      axis.text.y = element_text(size=rel(text_size), hjust=1, color='black'),
      axis.text.x = element_text(size=rel(text_size), vjust=0.5, color='black'),
      axis.title.y=element_text(size=rel(text_size), hjust=1, color='black'),
      axis.title.x=element_text(size=rel(text_size), vjust=1, color='black'),
      legend.text =   element_text(size=rel(text_size), hjust=1, color='black'), 
      legend.title =   element_text(size=rel(text_size), hjust=1, color='black'), 
      panel.grid.minor = element_blank(), 
      panel.border = element_rect(color = "black", fill = NA, size = 0.8),
      panel.background = element_blank(),
    ) 

ggsave(filename = paste0("search_labels/", "proj_k_5", ".pdf"), plot = p, width = 10, height = 7, dpi = 300)
```


```{r}
library(scatterplot3d)
scatterplot3d(factors_mofa[,1], factors_mofa[,2], factors_mofa[,3], color = colors[factor(km_res$cluster)], pch = 19)

colors <- c("1" = "blue", "2" = "red", "3" = "green", "4"="orange", "5" = "purple")


```

```{r}
library(plotly)
plot_ly(x = ~factors_mofa[,1], y = ~factors_mofa[,2], z = ~factors_mofa[,3], 
        color = factor(km_res$cluster), 
        colors = c("blue", "red", "green","orange", "purple"),
        type = "scatter3d", mode = "markers")

```


```{r}


plot_data <- data.frame(
  Time = survival_omics[["Time"]],
  Status = survival_omics[["Status"]],
  Cluster = factor(km_res$cluster,levels = 1:5, 
                                  labels = paste("Cluster", 1:5))
)


surv_obj <- Surv(time = plot_data$Time, event = plot_data$Status)
fit <- survfit(surv_obj ~ Cluster, data = plot_data)

names(fit$strata) <- gsub("Cluster=", "", names(fit$strata))

p <- ggsurvplot(
  fit,
  data = plot_data,
  risk.table = F,
  pval = F,
  palette = "Dark2",
  legend.title = "",
  xlab = "Time (days)",
  ylab = "Survival Probability"
)

p <- p$plot + theme(
      axis.text.y = element_text(size=rel(text_size), hjust=1, color='black'),
      axis.text.x = element_text(size=rel(text_size), vjust=0.5, color='black'),
      axis.title.y=element_text(size=rel(text_size), hjust=1, color='black'),
      axis.title.x=element_text(size=rel(text_size), vjust=1, color='black'),
      legend.text = element_text(size = rel(text_size), color = 'black'),
      legend.title = element_text(size = rel(text_size), color = 'black'),
      panel.grid.minor = element_blank(), 
      panel.border = element_rect(color = "black", fill = NA, size = 0.8),
      panel.background = element_blank(),
    )

ggsave(filename = paste0("search_labels/", "surv_curves_k_5", ".pdf"), plot = p, width = 10, height = 7, dpi = 300)
```



# Table with some information on the clusters and ground-truth labels

```{r}
metadata_clust <- metadata

metadata_clust$Cluster <- factor(km_res$cluster,levels = 1:5, 
                                  labels = paste("Cluster", 1:5))
metadata_clust$IDH1 <- omics.list$Mutations[, "IDH1"]
metadata_clust$EGFR <- omics.list$Mutations[, "EGFR"]
metadata_clust$PTEN <- omics.list$Mutations[, "PTEN"]
metadata_clust$TP53 <- omics.list$Mutations[, "TP53"]
metadata_clust$PIK3CA <- omics.list$Mutations[, "PIK3CA"]

cluster_summary <- metadata_clust %>%
  group_by(Cluster) %>%
  summarise(
    n = n(),
    AGE = round(mean(Age, na.rm = TRUE), 2),
    MALE = round(mean(Gender == "Male", na.rm = TRUE) * 100,2),
    OLIGO = sum(Subtype == "OLIGO", na.rm = TRUE),
    ASTRO = sum(Subtype == "ASTRO", na.rm = TRUE),
    GBM = sum(Subtype == "GBM", na.rm = TRUE),
    DEAD = round(sum(Vital_Status == "Dead", na.rm = TRUE)/n * 100,2),
    IDH1 = round(sum(IDH1 == 1, na.rm = TRUE)/n * 100,2),
    PTEN = round(sum(PTEN == 1, na.rm = TRUE)/n * 100,2),
    EGFR = round(sum(EGFR == 1, na.rm = TRUE)/n * 100,2),
    TP53 = round(sum(TP53 == 1, na.rm = TRUE)/n * 100,2),
    PIK3CA = round(sum(PIK3CA == 1, na.rm = TRUE)/n * 100,2)
  )

cluster_summary
write.csv(cluster_summary, "search_labels/info_clust5.csv", row.names = TRUE)
```

The samples in cluster 4, which histological type do they have?
```{r}
classification <- as.matrix(read.csv("DataSets/SIMPLIFIED_CLASSIFICATION_TCGA_2016_2021.csv", row.names = 1))
missing_samples <- setdiff(rownames(omics.list$Mutations),rownames(classification))
new_entries <- matrix("unclassified", 
                      nrow = length(missing_samples), 
                      ncol = ncol(classification),
                      dimnames = list(missing_samples, colnames(classification)))


classification <- rbind(classification, new_entries)


classification_1 <- classification[names(km_res$cluster)[km_res$cluster == 1], ]
classification_2 <- classification[names(km_res$cluster)[km_res$cluster == 2], ]
classification_3 <- classification[names(km_res$cluster)[km_res$cluster == 3], ]
classification_4 <- classification[names(km_res$cluster)[km_res$cluster == 4], ]
classification_5 <- classification[names(km_res$cluster)[km_res$cluster == 5], ]
```

```{r}
table(classification_1[, "TCGA.histological.type"]) # mostly untreated primary (de novo) gbm
table(classification_3[, "TCGA.histological.type"]) # mostly astrocytoma  and some untreated primary (de novo) gbm


table(classification_4[, "TCGA.histological.type"])

table(classification_2[, "TCGA.histological.type"]) 
table(classification_5[, "TCGA.histological.type"])


```

Let's re-label:
Cluster 1: GBM_1
Cluster 3: GBM_3
Cluster 2: ASTRO_2
Cluster 5: OLIGO_5
Cluster 4: MIX_LGG_4

```{r}
clust_labeled <- as.matrix(km_res$cluster)


new_labels <- c(
  "1" = "GBM_1",
  "2" = "ASTRO_2",
  "3" = "GBM_3",
  "4" = "MIX_LGG_4",
  "5" = "OLIGO_5"
)


clust_labeled_named <- new_labels[as.character(clust_labeled[, 1])]
names(clust_labeled_named) <- rownames(clust_labeled)

clust_labeled_named <- as.vector(clust_labeled_named)
```



# Repeat DGE  based on these groups 


```{r}
library("DESeq2")
library("edgeR")


mrna <- as.matrix(read.csv("DataSets/original_assays/assay_rna_coding.csv", row.names = 1))

mrna <- t(mrna)

keep <- rowSums(cpm(mrna) > 1) >= 5
mrna <- mrna[keep,]

mrna <- DGEList(counts = mrna)
mrna <- normLibSizes(mrna, method = "TMM")

patients <- rownames(omics.list$Mutations)

# only keep the 318 patients
mrna$samples <- mrna$samples[patients, ]
mrna$counts <- mrna$counts[,patients]

mrna$samples$group <- clust_labeled_named

group <- factor(clust_labeled_named)

# design matrix
design <- model.matrix(~ 0 + group, data = mrna$samples)
colnames(design) <- levels(group)
design


# dispersion
mrna <- estimateDisp(mrna, design, robust=TRUE)
mrna$common.dispersion


# fit 
fit <- glmQLFit(mrna, design, robust=TRUE)
head(fit$coefficients)
```


```{r}
con_1_2 <- makeContrasts(GBM_1 - ASTRO_2, levels=design)
con_1_3 <- makeContrasts(GBM_1 - GBM_3, levels=design)
con_1_4 <- makeContrasts(GBM_1  - MIX_LGG_4, levels=design)
con_1_5 <- makeContrasts(GBM_1  - OLIGO_5, levels=design)

con_2_3 <- makeContrasts(ASTRO_2 - GBM_3, levels=design)
con_2_4 <- makeContrasts(ASTRO_2 - MIX_LGG_4, levels=design)
con_2_5 <- makeContrasts(ASTRO_2  - OLIGO_5, levels=design)

con_3_4 <- makeContrasts(GBM_3 - MIX_LGG_4, levels=design)
con_3_5 <- makeContrasts(GBM_3 -  OLIGO_5, levels=design)

con_4_5 <- makeContrasts(MIX_LGG_4 - OLIGO_5, levels=design)




fit_1_2 <- glmQLFTest(fit, contrast=con_1_2)
fit_1_3 <- glmQLFTest(fit, contrast=con_1_3)
fit_1_4 <- glmQLFTest(fit, contrast=con_1_4)
fit_1_5 <- glmQLFTest(fit, contrast=con_1_5)

fit_2_3 <- glmQLFTest(fit, contrast=con_2_3)
fit_2_4 <- glmQLFTest(fit, contrast=con_2_4)
fit_2_5 <- glmQLFTest(fit, contrast=con_2_5)

fit_3_4 <- glmQLFTest(fit, contrast=con_3_4)
fit_3_5 <- glmQLFTest(fit, contrast=con_3_5)

fit_4_5 <- glmQLFTest(fit, contrast=con_4_5)

```
 

 
 
Factor 1 genes
```{r}
fac_1_de <- matrix(data = NA, nrow = 5, ncol = 5, dimnames = list(new_labels, new_labels))

for (i in 1:5){
  for (j in (i+1):5){
    contrast_str <- paste0(colnames(fac_1_de)[j], " - ", colnames(fac_1_de)[i])

    contrast_matrix <- makeContrasts(contrasts = contrast_str, levels = design)

    fit_rest <- glmQLFTest(fit, contrast=contrast_matrix)
    fit_rest <- fit_rest$table[mrna_sel_f1_, ]
    de_genes <- rownames(fit_rest)[abs(fit_rest$logFC) > 1 &
      fit_rest$PValue <= 0.05
    ]

    fac_1_de[i, j] <- length(de_genes)
  }
}
fac_1_de[lower.tri(fac_1_de)] <- t(fac_1_de)[lower.tri(fac_1_de)]

```


```{r}
fac_2_de <- matrix(data = NA, nrow = 5, ncol = 5, dimnames = list(new_labels, new_labels))

for (i in 1:5){
  for (j in (i+1):5){
    contrast_str <- paste0(colnames(fac_2_de)[j], " - ", colnames(fac_2_de)[i])

    contrast_matrix <- makeContrasts(contrasts = contrast_str, levels = design)

    fit_rest <- glmQLFTest(fit, contrast=contrast_matrix)
    fit_rest <- fit_rest$table[mrna_sel_f2_, ]
    de_genes <- rownames(fit_rest)[abs(fit_rest$logFC) > 1 &
      fit_rest$PValue <= 0.05
    ]

    fac_2_de[i, j] <- length(de_genes)
  }
}
fac_2_de[lower.tri(fac_2_de)] <- t(fac_2_de)[lower.tri(fac_2_de)]
```

```{r}
fac_3_de <- matrix(data = NA, nrow = 5, ncol = 5, dimnames = list(new_labels, new_labels))

for (i in 1:5){
  for (j in (i+1):5){
    contrast_str <- paste0(colnames(fac_3_de)[j], " - ", colnames(fac_3_de)[i])

    contrast_matrix <- makeContrasts(contrasts = contrast_str, levels = design)

    fit_rest <- glmQLFTest(fit, contrast=contrast_matrix)
    fit_rest <- fit_rest$table[mrna_sel_f3_, ]
    de_genes <- rownames(fit_rest)[abs(fit_rest$logFC) > 1 &
      fit_rest$PValue <= 0.05
    ]

    fac_3_de[i, j] <- length(de_genes)
  }
}
fac_3_de[lower.tri(fac_3_de)] <- t(fac_3_de)[lower.tri(fac_3_de)]
```

```{r}
fac_1_de
fac_2_de # this well separates the mix_lgg from all of the others - the question remains! 
fac_3_de # 
```
- GBM_1 seems more agressive than GBM_3 since: worst survival, older patients, histologicla type of mostly untreated primary (de novo) gbm;  they differ in terms of expression of factor 2 genes (they are up-regulated in GBM-3)
  is GBM-3 the neural subtype? this genes are related to nueral system, one of the genes were reprted as being this subtype in 
  who-gbm4subtypes


- Factor 1 genes separates the GBM groups from the others
- GBM_3 is more similar to LGG group as indicated by factor 2 table (the one that is astro or oligo, because not the mixed lgg)

- Factor 2 gens are crucial for MIX_LGG_4 to disintguish from all groups, this is actually a profile, less agresive than ASTRO_2 but more agressive than OLIGO_5; they are up-regulated compared to all of the groups.

are factor 2 genes agressive in oligo and less agressive in astro? 




```{r}
# MIX_LGG_4 -1, GBM_1 1
fit_1_4$table[mrna_sel_f2_, ] #logFC negative: genes up regulated in mix-lgg

# MIX_LGG_4 -1, ASTRO_2 1
fit_2_4$table[mrna_sel_f2_, ] #logFC negative: genes up regulated in mix-lgg

# MIX_LGG_4 -1, GBM_3 1
fit_3_4$table[mrna_sel_f2_, ] #logFC negative: genes up regulated in mix-lgg


# MIX_LGG_4 1, OLIGO_5 -1
fit_4_5$table[mrna_sel_f2_, ] #logFC positive : genes up regulated in mix-lgg



fit_1_3$table[mrna_sel_f2_, ]
con_1_3
```


- factor 3 genes may be common to GBM groups and astro. but well separates astro from oligo as seen before

## DMP
 
```{r}
library("ChAMP")

dna <- as.matrix(read.csv("DataSets/processed_assays/assay_methylation.csv", row.names = 1))
dna <-  dna[rownames(omics.list$Mutations), ]


dna_beta <- t(M2B(dna))

unique(clust_labeled_named)

```


```{r}
fac_1_de_dna <- matrix(data = NA, nrow = 5, ncol = 5, dimnames = list(new_labels, new_labels))

for (i in 1:5){
  for (j in (i+1):5){
    contrast_str <- c(colnames(fac_1_de_dna)[j], colnames(fac_1_de_dna)[i])

    res <- champ.DMP(beta = dna_beta,
                              pheno = clust_labeled_named,
                              compare.group = contrast_str,
                              adjPVal = 0.05 ,
                              adjust.method = "BH",
                              arraytype = "450K")
    res <- res[[1]]
    res <- res[methy_sel_f1, ]
    
    de_probes <- rownames(res)[abs(res$logFC) > 0.2 &
      res$adj.P.Val <= 0.05
    ]

    fac_1_de_dna[i, j] <- length(de_probes)
  }
}
fac_1_de_dna[lower.tri(fac_1_de_dna)] <- t(fac_1_de_dna)[lower.tri(fac_1_de_dna)]
```



```{r}
fac_3_de_dna <- matrix(data = NA, nrow = 5, ncol = 5, dimnames = list(new_labels, new_labels))

for (i in 1:5){
  for (j in (i+1):5){
    contrast_str <- c(colnames(fac_3_de_dna)[j], colnames(fac_3_de_dna)[i])

    res <- champ.DMP(beta = dna_beta,
                              pheno = clust_labeled_named,
                              compare.group = contrast_str,
                              adjPVal = 0.05 ,
                              adjust.method = "BH",
                              arraytype = "450K")
    res <- res[[1]]
    res <- res[methy_sel_f3, ]
    
    de_probes <- rownames(res)[abs(res$logFC) > 0.2 &
      res$adj.P.Val <= 0.05
    ]

    fac_3_de_dna[i, j] <- length(de_probes)
  }
}
fac_3_de_dna[lower.tri(fac_3_de_dna)] <- t(fac_3_de_dna)[lower.tri(fac_3_de_dna)]
```

 

## K = 3

# Plot the projection based on that assignment + Survival Curves of each cluster

```{r}

k2 <- 3

set.seed(64)
km_res2 <- kmeans(factors_mofa[,c(1,2,3)], centers = k2, nstart = 10)

```

```{r}

plot_proj_data <- data.frame(x = factors_mofa[,1], y  = factors_mofa[,3], color_by = factor(km_res2$cluster))

ggplot(plot_proj_data , aes(x=.data$x, y=.data$y, color=.data$color_by)) + 
    geom_point(size=3.0) +
    theme(
      axis.text.y = element_text(size=rel(text_size), hjust=1, color='black'),
      axis.text.x = element_text(size=rel(text_size), vjust=0.5, color='black'),
      axis.title.y= element_text(size=rel(text_size), hjust=1, color='black'),
      axis.title.x=element_text(size=rel(text_size), vjust=1, color='black'),
      legend.text = element_text(size = rel(text_size), color = 'black'),
      legend.title = element_text(size = rel(text_size), color = 'black'),
     panel.grid.minor = element_blank(), 
    panel.border = element_rect(color = "black", fill = NA, size = 0.8))


```



```{r}



plot_data <- data.frame(
  Time = survival_omics[["Time"]],
  Status = survival_omics[["Status"]],
  Cluster = factor(km_res2$cluster,levels = 1:3, 
                                  labels = paste("Cluster", 1:3))
)


surv_obj <- Surv(time = plot_data$Time, event = plot_data$Status)
fit <- survfit(surv_obj ~ Cluster, data = plot_data)

names(fit$strata) <- gsub("Cluster=", "", names(fit$strata))

p <- ggsurvplot(
  fit,
  data = plot_data,
  risk.table = F,
  pval = F,
  palette = "Dark2",
  legend.title = "",
  xlab = "Time (days)",
  ylab = "Survival Probability"
)

p <- p$plot + theme(
      axis.text.y = element_text(size=rel(text_size), hjust=1, color='black'),
      axis.text.x = element_text(size=rel(text_size), vjust=0.5, color='black'),
      axis.title.y=element_text(size=rel(text_size), hjust=1, color='black'),
      axis.title.x=element_text(size=rel(text_size), vjust=1, color='black'),
      legend.text = element_text(size = rel(text_size), color = 'black'),
      legend.title = element_text(size = rel(text_size), color = 'black'),
      panel.grid.minor = element_blank(), 
      panel.border = element_rect(color = "black", fill = NA, size = 0.8),
      panel.background = element_blank(),
    )

ggsave(filename = paste0("search_labels/", "surv_curves_k_3", ".pdf"), plot = p, width = 10, height = 7, dpi = 300)

```


```{r}
metadata_clust2 <- metadata

metadata_clust2$Cluster <- factor(km_res2$cluster)
metadata_clust2$IDH1 <- omics.list$Mutations[, "IDH1"]
metadata_clust2$EGFR <- omics.list$Mutations[, "EGFR"]
metadata_clust2$PTEN <- omics.list$Mutations[, "PTEN"]
metadata_clust2$TP53 <- omics.list$Mutations[, "TP53"]

cluster_summary2 <- metadata_clust2 %>%
  group_by(Cluster) %>%
  summarise(
    n = n(),
    AGE = mean(Age, na.rm = TRUE),
    MALE = mean(Gender == "Male", na.rm = TRUE) * 100,
    OLIGO = sum(Subtype == "OLIGO", na.rm = TRUE),
    ASTRO = sum(Subtype == "ASTRO", na.rm = TRUE),
    GBM = sum(Subtype == "GBM", na.rm = TRUE),
    DEAD = sum(Vital_Status == 1, na.rm = TRUE)/n * 100,
    IDH1 = sum(IDH1 == 1, na.rm = TRUE)/n * 100,
    PTEN = sum(PTEN == 1, na.rm = TRUE)/n * 100,
    EGFR = sum(EGFR == 1, na.rm = TRUE)/n * 100,
    TP53 = sum(TP53 == 1, na.rm = TRUE)/n * 100
  )

cluster_summary2
```

```{r}
table(classification[names(km_res2$cluster)[km_res2$cluster == 3], "TCGA.histological.type"])
table(classification[names(km_res2$cluster)[km_res2$cluster == 1], "TCGA.histological.type"])
```








